We need to beat namespotting, no idea how

### 2023-02-01 paperSubmission/paperSubmission1.tex
final version of the paper in zip

\begin{abstract}
Abstract goes here.

\end{abstract}

@@

\begin{keyword}
 keyword \sep keyword

@@

@@

The term counter-speech refers to the use of means of communication to counteract hate speech or other forms of harmful or extremist speech. It is a form of counter-narrative that seeks to challenge and undermine the messages of hate and extremism, and to promote alternative, positive, and inclusive narratives. It can be broadly defined as citizens’ response to hate speech with the purpose of stopping it and reducing its consequences \citep{leader2016dangerous, garland2020countering}. Hate speech is a type of speech used with the intention to insult, offend, or intimidate a person based on their identity (such as race, religion, sexual orientation, national origin, or disability) \citep{dictionary2002merriam}. This type of speech is often used to spread hateful or discriminatory ideas, both offline and online, and can take many forms, such as name-calling, threats, or slurs. Social media platforms often recommend their users to employ counter speech as a way to stop hate speech from spreading, conveniently handing the problem out to dedicated users to avoid creating new technologies or investing in manual moderation \citep{schieb2016governing}.

\citet{benesch2016considerations} defined successful counter-speech in two ways. The first type has a favorable impact on the hateful user, for example by shifting their discourse or beliefs, and the second one impacts the whole audience of counter-speech, by shifting the norm of the conversation.  \citep{rieger2018hate} proposed a three-layered classification of counter-speech: primary prevention, where counter-voices are being disseminated to the general public in order to educate and promote civic behavior; secondary prevention, where counter-speech is being directed at specific groups, for example, those more vulnerable to harmful effects of online hate, and tertiary prevention where counter-voices respond directly to those spreading hate. Our responses fulfilled all these functions, as they were publicly visible comments sent as direct responses to hateful users.

When it comes to the effectiveness of online counter-speech there is just a handful of large-scale classification studies. One of them used a special situation in Germany, where groups that label themselves as hateful or counter-speaking talked about immigration and elections. Results based on analysis of 130,000 Twitter conversations suggest that counter speech can help make discussions less divisive, encourage more counter-speech and lead to fewer hateful responses \citep{garland2020countering}.

Another study by \citet{ziegele2018journalistic} looked at almost 10,000 comments on the Facebook pages of 15 news organizations in Germany to investigate the patterns, determinants, and potential effects of interactive moderation. The authors found that the moderation style of initial comments was related to the presence of incivility in users' subsequent reply comments. Specifically, a sociable moderation style was associated with a decrease in incivility in reply comments, while a regulative style was associated with an increase in incivility. This shows that the style of counter responses can also determine their effectiveness.

Moreover, the work by \citet{ziems2020racism} looked at the dynamics of the spread of racial hate during a pandemic and the role of counter speech in mitigating this spread. Using a hand-labeled data set of 3,355 tweets, they trained a text classifier to identify hate and counter speech tweets and conducted a longitudinal analysis of tweets and users. Interestingly, hateful and counter speech users interacted and engaged extensively with one another, rather than staying in isolated, polarized communities. Results revealed that Twitter users were more likely to become hateful after being exposed to hateful content and that counter speech messages may discourage users from becoming hateful, suggesting that consistent counter speech could be a potential solution for effects of hate on social media platforms.

 \citep{xie2011social} examined the influence of committed minorities, and looked at how a small group of people with an opinion different from the majority can quickly change the whole group's opinion. They found that if more than 10 percent of the group expresses another viewpoint, there is a dramatic decrease in the time taken for the entire population to adopt it. In other words, a relatively small group of committed individuals can have a significant impact on the opinion of the majority.

Similarly, \citep{schieb2016governing} used a computer simulation to study whether counter-speech can curb hate speech on Facebook. They found that even a small group of users speaking out against hate can influence a much larger group, as long as there are enough people in the audience who do not hold extreme opinions. They also emphasized the potential negative effects of counter-voices, especially when they do not hold to the same standards of civic discussion but instead act in an authoritative or mocking way towards those expressing extremist ideologies.

Although some studies suggest censorship and elimination of hateful content to be a more effective strategy of toxicity reduction \citep{alvarez2018normative}, it also also collide with civil liberties and diffuse rather than reduce hate \citep{chandrasekharan2017you}. Therefore counter-speech seems to be a golden means for reducing antisocial behavior while preserving civil liberties.

In our study, we have employed counter-speaking strategies to see whether they can be applied to a slightly different phenomenon, namely personal attacks. Personal attacks are instances of speech that are intended to hurt, offend, or belittle a specific person or group. This can include name-calling, insults, or other types of derogatory language that is directed at an individual or group \citep{bilewicz2021artificial}. While personal attacks and hate speech can sometimes overlap, the main difference between the two is that personal attacks are focused on an individual or a specific group of people, whereas hate speech is aimed at an entire group of people based on their identity.

% we have previously studied the effectiveness of counter-speech in lowering the use of personal attacks on Reddit using an automatic intervention system  \citep{bilewicz2021artificial}.

%In this follow-up study we wanted to see whether

%\citet{benesch2016considerations} recommend counter-speech strategies based on empathy and affiliation, employing humor, warning of consequences of such behavior, labeling and shaming.
%Although censorship and hateful content elimination has shown to be a more effective strategy of toxicity reduction @alvarez2018normative, it can also collide with civil liberties and diffuse rather than reduce hate @chandrasekharan2017you. Therefore counter-speech seems to be a golden mean for reducing antisocial behavior while preserving civil liberties. 
Although counter-speech can be an effective strategy for toxic discourse mitigation, it usually occurs as a spontaneous act, as it is determined by personal motivations of the counter-speaking party. For example, \citet{brauer2010descriptive} found that being personally implicated by a counter-normative behavior was a primary determinant of the bystander reaction.

Apart from that, the more public the space, the less likely it is that someone will step up, as studies on bystander interventions consistently show that presence of other people in critical situations reduces the probability of assistance from bystanders \citep{fischer2011bystander}. %It is much less likely to happen in natural settings then in a laboratory. @?

Moreover, even if a bystander decides to react, they will not necessarily choose the most effective strategy, which can lead to potentially undesirable side-effects. For instance,  counter-voices that do not hold up to civic standards themselves but act in an authoritative manner or mock those falling for extremist ideologies, can actually strengthen the underlying attitudes \citep{ziegele2018journalistic, schieb2016governing, legault2011ironic}.
%empathy here somewhere

The issue of low bystander reactivity has been addressed in our study by engaging volunteers personally committed to lowering online violence, and creating an opportunity structure for quick response, by real-time notifications about the occurrences of personal attacks that they should respond to (see however our section on volunteer engagement for a discussion of difficulties with this strategy).

To maximize the potential impact of our interventions and to minimize  accidental instigation of more verbal aggression, we decided to educate the volunteers on scientifically grounded strategies for normative influence and persuasive messaging.  Let us elaborate.

ocial norms are agreed to be powerful determinants of human behavior. 
According to theories of norm compliance people tend to comply with the norms they are exposed to \citep{morris2018common, cialdini1990focus}.

Multiple studies have found that perceptions of what others are doing cause us to behave similarly. Norms can be applied to exert persuasive impact, for example, to solve a social problem such as littering or wasting resources \citep{cialdini2006managing, cialdini1990focus, goldstein2008room}.

%Researchers have repeatedly found that perceptions of what others are doing influences subjects to behave similarly, for example when it comes to choosing consumer product (venkatesen, 1966), or looking at the sky (milgram, bichman, berkowitz, 1969).

%Reasons why social norms have such strong influence on peoples decisions, attitudes, and the way they express them are still disputed. %cytowanie

Some scholars propose that perceived social norms serve as shortcuts helping us quickly adapt to new environments \citep{cialdini1990focus}. Others say that we are only motivated by norms we have internalized during socialization, so the normative influence can only be exerted if the norm resonates with our internal value system \citep{morris2018common}.  %cytowania

The Focus Theory of Normative Conduct \citep{kallgren2000focus} is a  theoretical psychological  framework used for understanding best conditions and practices for exerting normative influence. Its main assumption is that norms that are focal in attention are likely to influence behavior. One important distinction made within this approach is between two types of norms: a descriptive norm refers to what is typically done, by showing an example of normal behavior in a given situation, and an injunctive or prescriptive norm refers to what is commonly approved, such as the rules of proper conduct and moral beliefs \citep{cialdini1990focus}.

A descriptive norm motivates compliance by providing evidence about what is likely to be an effective and adaptive action, while an injunctive norm motivates by informing the agent which actions will be approved and which will result in social sanctions \citep{yanowitzky2006communication, cialdini2004social, kallgren2000focus}.

%This has important implications for normative influence practitioners, as the same behavior can be prescriptively normative and descriptively not normative, and vice versa. In other words, something may be the right thing to do but few people actually do it  \citep{brauer2010descriptive}.
%truism. say more

The positive influence of injunctive and descriptive norms on unwanted social behavior was first presented in a study on littering in public spaces conducted by \citet{cialdini1990focus}. The study subjects were exposed to either descriptive or injunctive anti-littering norms, and the impact on the littering behavior was observed. In the descriptive condition, the subjects were exposed to a clean or littered environment, serving as proof of choices made by others. In the injunctive condition, the participants received messages stating the anti-littering norm.

Injunctive messages directly asking not to litter had significantly reduced littering. The subjects littered more in a highly littered environment and littered less in a clean environment, although these effects were not statistically significant.\footnote{It is also worth noticing that  seeing litter in an otherwise clean environment could have also engaged an injunctive norm, as it might have reminded subjects about societal objections to littering.}

%Some of these results were replicated \citep {berquisr2021replicating} \citep {berger2016norm} %say more

A descriptive norm can be also evoked by providing written information about what others usually do. \citet{goldstein2008room} distributed flyers with normative messages that contained descriptive information about reuse of hotel towels (“the majority of guests reuse their towels”). This strategy has been proven superior to traditional appeals focusing on environmental protection. Moreover, normative appeals were most effective when the described behavior occurred in the same location (“in this hotel  room”), and so referred to, so called, local norms.

The information about how others think and behave can also have an impact on prejudice and its expression. For example, positive information about the racial beliefs of others produced stereotype change in message recipients \citep{stangor2001changing}, and public expression of prejudice, and reactions to hostile jokes were highly correlated with social approval of these expressions \citep{crandall2002social}.

When it comes to the online realm, group norm promotion and social sanctioning of online racist harassment resulted in the reduction of racist tweets (although only in the condition in which a white male with high followers was approaching another white male). In other words, sending direct, counter-speech messages to users with similar identities,  resulted in a decrease in their use of hate speech on Twitter \citep{munger2017tweetment}. Moreover, in a study conducted by \citet{bilewicz2021artificial}, a bot disguised as a Reddit user, equipped with normative and empathetic interventions, significantly decreased the number of personal attacks generated in this forum. In addition, types of statements made by people in online political discussions influenced the way others expressed their opinions \citep{price2006normative}, and the expression of online hate-speech in  \citep{alvarez2018normative} was strongly dependant on the amount of hate-speech already present in a given environment.

The context of online interactions can also contribute to the spread of counter-normative behaviors. For example, negative mood and prior exposition to troll content was the strongest predictor for trolling behavior in \citep{cheng2017anyone}.

In line with the above discussion, for the purpose of this study we have distinguished two types of responses to personal attacks: "empathetic" and "normative". The empathetic approach is aimed at countering the so-called desensitization to violence. It is a very dangerous process, finally leading to the dehumanization of stigmatized groups.  
 Being frequently exposed to online hate speech builds tolerance to the phenomenon and as a result, we no longer see such statements as offensive, which lowers our sympathy for the victims and increases prejudice. By sending empathetic interventions we hoped to evoke compassion and self-reflection, and  to humanize the person on the other side of the screen \citep{bilewicz2021artificial}.

The normative approach stems from the  already discussed psychological research of normative social influence, stating that in ambiguous situations, people tend to rely on social cues, to determine what kinds of behaviors will be accepted. Taking from these studies, we refer to two types of norms: injunctive norms---describing what the members of this community are supposed to do and what is approved or disapproved by others (laws, rules, code of conduct, savoir-vivre, etc.) and descriptive norms---what the members of this community actually do \citep{cialdini_focus_1991, cialdini2004social, cialdini2006managing}.

\section{Collective Intelligence and  moderation}\label{collective-intelligence-approach-to-counter-speech}

@@

Instead of using a negative motivation system, the assumption was to test a positive one—convincing verbally violent users to refrain from using cyberviolence based on civic education—peer-pressure regulation and experiential learning of a positive set of norms and empathy. Algorithms developed for the detection of personal attacks were used to monitor the activity of experimental groups and notify volunteers about all attacks generated by its members. Volunteers, after receiving a notification on Slack, could then react with a proper intervention. Such an approach served as a distributed bottom-up voluntary model of moderation based on collective intelligence—utilizing human + machine intelligence to develop solutions to social challenges.

Normative interventions can refer either to general social norms of civility and respect, community standards, a particular subreddit rule, or a descriptive norm among others. Empathetic interventions can refer to the emotional state of the recipient or the sender of the attack or both. They are designed to evoke an empathetic response. Here are some examples of interventions that were used during the experiment: 
\begin{itemize}
  \item \emph{Insulting others is against Reddit's policy.} (normative)
  \item \emph{Hey there, we do not call each other like that here.} (normative)
  \item \emph{There is a human being on the other side who might be hurt by your words.} (empathetic)
  \item \emph{I see you are frustrated but remember the human.} (empathetic)
\end{itemize}

@@

%the reasoning behind H2
Although empathy has been found to be helpful in the reduction of aggressive behavior \citep{bjorkqvist2000social}, and the improvement of intergroup relations \citep{stephan1999role}, there are also important arguments and evidence concerning its limits. Specifically, Paul Bloom in his book \emph{}Against Empathy: The Case for Rational Compassion} \citep{bloom2017against}, stresses that empathy can have a very narrow focus and can be prone to biases. For instance, it’s easier to be empathetic towards those who are close or similar to us. Since conversations on Reddit are anonymous and usually occur between strangers, we expected  empathetic interventions to  be less effective than normative interventions in curbing personal attacks.

@@

\textit{Personal attacks} are defined as any
kind of abusive remark made in relation to a person
(\textit{ad hominem}) rather than to the content of the argument
expressed by that person in a discussion. The definition of `personal
attack' subsumes the use of specific terms which compare other people to
animals or objects or making nasty insinuations without providing
evidence. Three examples of typical personal attacks are as follows:

\begin{itemize} 
\item \emph{You are legit mentally retarded homie.}
\item \emph {Eat a bag of dicks, fuckstick.}
\item \emph {Fuck off with your sensitivity you douche.} 
\end{itemize}

@@

We conducted a 6-months field experiment in a digital setting conducted on a popular Q\&A and news forum, Reddit (\url{www.reddit.com}). We formed treatment and control groups based on three main criteria:

@@

%update when done

>>>>

\begin{abstract}
During a six-month experiment conducted on Reddit, we studied the impact  of  counter-speech interventions against personal attacks on the number of personal attacks sent by  440  users regularly attacking others. We used two types of interventions, normative---which referred to social norms, and empathetic---which referred to emotions and encouraged perspective-taking. We employed a  collective intelligence approach---the collaboration between human and machine intelligence. Artificial Intelligence was used to detect verbal aggression and notify human volunteers, who then performed the interventions. We analyzed the data from three perspectives. We used time series models (1) of the short-term impact of individual interventions, and of (2) the cumulative impact of interventions received as the experiment progressed.  We also (3) used aggregated  data for a long-term before/after analysis.  The short-term effect of interventions is damaging: users tend to be on average around 26\% more aggressive the next day, but the effect does not last beyond two days. The cumulative effect of interventions is helpful: each intervention (up to around 8-10 total, the effectiveness of more interventions tends to be lower) decreases daily aggression by 4\% on average and the effects accumulate and  balance out the short-term effect in the long run. The effectiveness of normative interventions seems overall higher, except for the less aggressive offenders, for which empathetic interventions might be equally or more useful. 
\end{abstract}

@@

\begin{keyword}
counter-speech\sep artificial intelligence\sep collective intelligence\sep  peaceful interventions\sep  Reddit\sep   no i nie wiem jeszcze który do przemocy słownej: online aggression\sep  online abuse\sep  personal attacks

@@

\section{Introduction}\label{introduction}

Much effort has been made to tackle the problem of verbal aggression and harassment online, but looking at various reports and surveys \citep{Zachary_hate, sorrentino2019epidemiology, vogels_state_2021}, it remains a common hindrance for people who engage with social media in their everyday lives. The problem was exacerbated during the COVID-19 pandemic, during which much of our social life moved to cyberspace. This shift was followed by an increase in cyberbullying attitudes and perpetration \citep{barlett2021comparing}, 90\% increase in public reports of illegal online content \citep{grant_2021}, including 114\% increase in non-consensual sharing of intimate images, 30\% increase in cyberbullying, as well as 40\% of increase in adults reporting online harassment. According to a report by company L1ght \citep{noauthor_l1ght_2020}, hate speech directed towards China and the Chinese went up by 900\% on Twitter. Gaming platforms were in the spotlight as well, with a 40\% increase in toxicity on Discord.\footnote{In both reports, the increase is reported as a relative change between the years 2019 and 2020, with no absolute indicators.}

Unfortunately, the growing need for even more efficient and proactive moderation was not followed by an increased capacity to execute it, forcing companies and policymakers to rethink the current model of moderation processes and workforce. Due to the COVID-19 restrictions, including social distancing, a lot of those serving the role of moderators had to be sent home \citep{bhattacharya} without the ability to work remotely because of the constraints affiliated with restrictive non-disclosure agreements (NDA), among others. Curtailing the moderators’ workforce was accompanied by more agency given to algorithms and AI-based moderation. As argued by \citet{gerrard2020covid19}, those changes are a serious safety red flag for all the users of online platforms. At the same time, looking at the prevalence of online violence and the scope of the problem, cooperation between humans and algorithms seems to be inevitable. The question is how to execute such a partnership, taking into consideration the technological constraints and human limitations, and the strong suits of both.

The matter in question is of great significance, as online violence (including cyberbullying, sexual harassment, hate speech, or any other type of abuse online) poses a great risk to individuals and groups alike with its harrowing consequences. On an individual level, online abuse was found to correlate with anxiety, lower self-esteem, feeling of hopelessness, isolation, substance abuse, decreased academic performance, and in more extreme cases self-harm, suicidal ideation, or suicide \citep{yang2021consequences, hinduja2010bullying, alhujailli2020affective, kowalski2013psychological}. On a group level, it has been linked to group polarisation and radicalization \citep{lee2002persuasive, williams2019hatred}. Online hatred running rampant on social media platforms can incite atrocities such as genocide---The Rohingya genocide was preceded by an increase in anti-Muslim rhetoric and a genocidal campaign occurring on Facebook \citep{fink2018dangerous}.
There is also research on how the prevalence of hate speech can lead to desensitization---a decreased ability to judge hateful statements as hateful, and in turn to an increase of prejudice \citep{soral2018exposure}, and of social acceptability of hate speech \citep{alvarez2018normative}, or increase of offline hate crime \citep{williams2020hate}. 
Online abuse is also a threat to the stability of online services and their communities. It can contribute to lower user engagement or churn \citep{urbaniak2022personal}, directly impacting the profitability of companies. In some countries, there are new regulations that introduce severe fines for companies not reacting to reported hate speech or other types of online harm \citep{tworek2021fighting, trengove2022critical}. In the past, there were also instances in which companies pulled their ads from online platforms, where violent speech was prevalent \citep{cnbc_2020}. On the more extreme side, some companies were sued or even shut down because of online abuse and suicides related to it \citep{newton_2013}. Cyberbullying and gun-related threats have been found to play a major role in YikYak decline and shutdown \citep{failory_2022}. Snapchat was sued by a mother of a teen who committed suicide after being bullied on a platform \citep{los_angeles_times_2021}, while Formspring closed down in 2013 after a series of suicides among its users \citep{seifert_2013}.

In this paper, we describe  a novel experiment in the collective intelligence paradigm designed to reduce the level of verbal aggression on the Q\&A forum Reddit. The collective intelligence approach in its underlying principles entails cooperation between humans and machines to solve a particular problem. Detection  algorithms were used to automatically detect verbal aggression in the form of personal attacks and to notify human volunteers, who then responded with counter-speech interventions aimed at the reduction of verbal aggression in  study  participants.

This paper is organized as follows: first, we will sketch the landscape of online content moderation and locate the techniques we used  within it. Second, we will outline the theoretical underpinnings of counter-speech interventions utilized in the experiment. Then, we describe the experiment itself. Having done this, we wrap up  with a  short discussion of  volunteer engagement in this experiment.

@@

The term counter-speech refers to the use of means of communication to counteract hate speech or other forms of harmful or extremist speech. It is a form of counter-narrative that seeks to challenge and undermine the messages of hate and extremism, and to promote alternative, positive, and inclusive narratives. It can be broadly defined as citizens’ response to hate speech aimed to stop it and reduce its consequences \citep{leader2016dangerous, garland2020countering}. Hate speech is a type of speech used to insult, offend, or intimidate a person based on their identity (such as race, religion, sexual orientation, national origin, or disability) \citep{dictionary2002merriam}. This type of speech is often used to spread hateful or discriminatory ideas, both offline and online, and can take many forms, such as name-calling, threats, or slurs. Social media platforms often recommend their users to employ counter speech as a way to stop hate speech from spreading, conveniently handing the problem out to dedicated users to avoid creating new technologies or investing in manual moderation \citep{schieb2016governing}.

\citet{benesch2016considerations} defined successful counter-speech in two ways. The first type has a favorable impact on the hateful user, for example by shifting their discourse or beliefs, and the second one impacts the whole audience of counter-speech, by shifting the norm of the conversation.  \citet{rieger2018hate} proposed a three-layered classification of counter-speech: primary prevention, where counter-voices are being disseminated to the general public to educate and promote civic behavior; secondary prevention, where counter-speech is being directed at specific groups, for example, those more vulnerable to harmful effects of online hate, and tertiary prevention where counter-voices respond directly to those spreading hate. Our responses fulfilled all these functions, as they were publicly visible comments sent as direct responses to hateful users.

When it comes to the effectiveness of online counter-speech there is just a handful of large-scale  studies, and even fewer of those who distinguish between different types thereof. One of them used a special situation in Germany, where groups that label themselves as hateful or counter-speaking talked about immigration and elections. Results based on analysis of 130,000 Twitter conversations suggest that counter-speech can help make discussions less divisive, encourage more counter-speech, and lead to fewer hateful responses \citep{garland2020countering}.

Another study by \citet{ziegele2018journalistic} looked at almost 10,000 comments on the Facebook pages of 15 news organizations in Germany to investigate the patterns, determinants, and potential effects of interactive moderation. The authors found that the moderation style of initial comments was related to the presence of incivility in users' subsequent reply comments. Specifically, a sociable moderation style was associated with a decrease in incivility in reply comments, while a regulative style was associated with an increase in incivility. This shows that the style of counter-responses can also determine their effectiveness.

Moreover,  \citet{ziems2020racism} looked at the dynamics of the spread of racial hate during a pandemic  and the role of counter-speech in mitigating this spread. Using a hand-labeled data set of 3,355 tweets, the authors trained a text classifier to identify hate and counter-speech tweets and conducted a longitudinal analysis of tweets and users. Interestingly, hateful and counter-speech users interacted and engaged extensively with one another, rather than staying in isolated, polarized communities. Results revealed that Twitter users were more likely to become hateful after being exposed to hateful content and that counter-speech messages may discourage users from becoming hateful, suggesting that consistent counter-speech could be a potential remedy for hate on social media platforms.

\citet{xie2011social} examined the influence of committed minorities and looked at how a small group of people with an opinion different from the majority can quickly change the whole group's opinion. They found that if more than 10 percent of the group expresses another viewpoint, there is a dramatic decrease in the time taken for the entire population to adopt it. In other words, a relatively small group of committed individuals can have a significant impact on the opinion of the majority.

Similarly, \citet{schieb2016governing} used a computer simulation to study whether counter-speech can curb hate speech on Facebook. They found that even a small group of users speaking out against hate can influence a much larger group, as long as there are enough people in the audience who do not hold extreme opinions. They also emphasized the potential negative effects of counter-voices, especially when they do not hold to the same standards of civic discussion but instead act in an authoritative or mocking way towards those expressing extremist ideologies.

Although some studies suggest censorship and elimination of hateful content to be a more effective strategy of toxicity reduction \citep{alvarez2018normative}, such methods also collide with civil liberties and diffuse rather than reduce hate \citep{chandrasekharan2017you}. Therefore, at least \emph{prima facie}, counter-speech seems to be a golden means for reducing antisocial behavior while preserving civil liberties.

In our study, we have applied counter-speech interventions to a slightly different phenomenon, namely personal attacks. Personal attacks are instances of speech that are intended to hurt, offend, or belittle a specific person or group. These can include name-calling, insults, or other types of derogatory language that is directed at an individual or group \citep{bilewicz2021artificial}. While personal attacks and hate speech can sometimes overlap, the main difference between the two is that personal attacks are focused on an individual or a specific group of people, whereas hate speech is aimed at an entire group of people based on their identity.

Although counter-speech can be an effective strategy for toxic discourse mitigation, it usually occurs as a spontaneous act, as it is determined by personal motivations of the counter-speaking party. For example, \citet{brauer2010descriptive} found that being personally implicated by a counter-normative behavior was a primary determinant of the bystander reaction. Apart from that, the more public the space, the less likely it is that someone will step up, as studies on bystander interventions consistently show that the presence of other people in critical situations reduces the probability of assistance from bystanders \citep{fischer2011bystander}. Moreover, even if a bystander decides to react, they will not necessarily choose the most effective strategy, which can lead to potentially undesirable side effects. 
For instance, counter-voices that do not hold up to civic standards themselves but act in an authoritative manner or mock those falling for extremist ideologies can actually strengthen the underlying attitudes \citep{ziegele2018journalistic, schieb2016governing, legault2011ironic}.

To maximize the potential impact of our interventions and to minimize accidental instigation of more verbal aggression, we decided to educate the volunteers on scientifically grounded strategies for normative influence and persuasive messaging. Drawing from research on social compliance and aggressive behavior, we have created two intervention categories---empathetic and normative---each based on a different mechanism of influence. We have used the distinction from our previous experiment where normative and empathetic interventions were automatically generated by a bot in response to personal attacks on Reddit. Both types of responses significantly decreased the prevalence of personal attacks, although the study had its limitations. For example, the interventions (in a form of public comments) were sent within certain subreddits, so we could not control if the users from one intervention group have not seen interventions from another group 
\citep{bilewicz2021artificial}. The current study addresses these limitations by employing a fully randomized experimental design and extending the study to the entire Reddit platform. Coming back to our example, this means that the probability that a subject encouters an intervention directed at someone else is much lower.

Another difference is that in the previous study, the interventions were generated automatically, and in this study replies were written and sent by human volunteers. We have engaged volunteers personally committed to lowering online violence and created an opportunity structure for quick response, by providing real-time notifications about personal attacks that should receive an intervention. We hoped that choosing intrinsically motivated volunteers, and providing them with guidance and AI assistance would solve the issue of bystander apathy (see however our section on volunteer engagement for a discussion of difficulties with this strategy).

\textbf{Empathetic interventions.} Empathetic interventions were thoughtful responses delivered in a gentle and straightforward manner,  empathizing with either the sender of the personal attack, the user who received it, or the both of them.  We hypothesized that presenting hostile individuals with empathy and kindness would reduce further attacks by interfering with their script of aggressive behavior. However, we had no way of verifying the direct effects of our interventions other than the change in the number of personal attacks generated by users.

According to literature, empathy can decrease the tendency to act in a hateful and derogatory manner 
\citep{zaki2014empathy}. Studies on desensitization to hate-speech against refugees conducted by \citet{bilewicz2016psychological} and \citet{soral2022role} found empathy to be an effective countermeasure to this phenomenon. Hate-speech desensitization is a lowered sensitivity to hate speech caused by frequent exposition to it.  As a result, the agent stops perceiving such speech as harmful, which in turn increases their prejudice towards the victims, and, in extreme cases, leads to the dehumanization of stigmatized groups.  Empathy-inducing interventions made subjects more sensitive to hateful content and more supportive of hate-speech prohibition.

Research on peer aggression has shown that both cognitive and affective empathy can reduce bullying behavior and increase the likelihood of a bystander helping a victim \citep{van2015empathy}. A study looking at predictors of either engaging in defending or remaining an outsider in bullying situations, found empathy to be a strong predictor of defending behaviors \citep{nickerson2008attachment}. Empathy was also investigated as a factor that can mitigate the negative effects of high social status on bullying behavior amongst 461 primary and secondary school students. The results confirmed that high levels of empathy suppressed involvement in bullying and moderated the relationship between bullying and social status
 \citep{caravita2009unique}.

Similar results were obtained regarding online aggression. For instance, online aggression towards celebrities was shown to be linked to low levels of empathy and high levels of moral disengagement 
\citep{ouvrein2018online}, and a survey of 2070 secondary school students found that cyberbullies had less empathy \citep{steffgen2011cyberbullies}. 
A study \citep{barlinska2018cyberbullying} looking at empathy and cyber-bystanding behavior among adolescents, found both affective and cognitive empathy to be positively related to prosocial cyber bystander behavior, with affective empathy being a stronger predictor than cognitive empathy. Moreover, the study investigating the bystander effect in cyberbullying has shown that factors such as empathic concern and immediate empathic response increased provided
 support \citep{machackova2015brief}.

 \citet{wright2021does} conducted a longitudinal analysis of the consequences of being a bystander to homophobic cyberbullying on a sample of 1067 adolescent students. According to their findings, witnessing homophobic cyberbullying led to increased perpetration of homophobic cyberbullying, but this effect was moderated by empathy and toxic online disinhibition. In other words, high empathy level was one of the factors preventing students from perpetrating online aggression.

Finally, \citet{chin2020empathy} investigated how conversational agents (such as chatbots and smart speakers) should respond to being verbally abused by their users. Specifically, they wanted to understand what kind of response style has the most positive impact on emotions that mitigate users’ aggressive behaviors. The authors hypothesized that empathetic responses addressing the harm caused by the users will reduce further abuse of the agent by eliciting feelings of guilt in the abusers. Indeed, the empathetic response style was most effective in reducing verbal abuse towards the conversational agent, although the feeling of guilt was not directly measured in this study.

\textbf{Normative interventions.} The normative interventions were based on the principle of normative influence. Their objective was to reduce personal attacks by lowering the perceived acceptability of such behavior in a particular online setting. The normative approach stems from the psychological research of normative social influence, stating that in ambiguous situations, people tend to rely on social cues, to determine what kinds of behaviors will be accepted.

Research has repeatedly shown that perceptions of what others are doing cause us to behave similarly. Some scholars propose that it is because perceived social norms serve as shortcuts helping us to quickly adapt to new environments \citep{cialdini1990focus}. Others say that we are only motivated by norms we have internalized during socialization, so the normative influence can only be exerted if the norm resonates with our internal value system \citep{morris2018common}. Either way, social norms are agreed to be powerful determinants of human behavior that can be applied to solve social problems, such as littering or resource waste \citep{cialdini2006managing, cialdini1990focus, goldstein2008room, morris2018common}.

One important distinction made within this approach is between two types of norms: 
descriptive norms that refer to what is typically done and are related to  how others behave, and prescriptive (or injunctive) norms that explicitly describe what is commonly approved, for example, the rules of proper conduct and moral beliefs \citep{cialdini1990focus}. In other words, a descriptive norm motivates compliance by providing evidence about what is likely to be effective and adaptive action, while a prescriptive  norm motivates by informing the agent which actions are approved and which will result in social sanctions \citep{yanowitzky2006communication, cialdini2004social, kallgren2000focus}.

In a study on littering in public spaces conducted by \citet{cialdini1990focus}, study subjects were exposed to either descriptive or injunctive anti-littering norms, and the impact on the littering behavior was observed. In the descriptive condition, the subjects were exposed to a clean or littered environment, serving as proof of choices made by others. In the injunctive condition, the participants received messages stating the anti-littering norm. Results have shown that prescriptive messages directly asking not to litter had significantly reduced littering, and the subjects littered more in a highly littered environment and littered less in a clean environment, although these effects were not statistically significant.\footnote{It is also worth noticing that seeing litter in an otherwise clean environment could have also engaged a prescriptive norm, as it might have reminded subjects about societal objections to littering.}

Descriptive norms can also be evoked by providing written information about what others usually do. \citet{goldstein2008room} distributed flyers with normative messages that contained descriptive information about the reuse of hotel towels (“the majority of guests reuse their towels”). This strategy has been proven superior to traditional appeals focusing on environmental protection. Moreover, normative appeals were most effective when the described behavior occurred in the same location (“in this hotel room”), and so referred to, so-called, local norms.

The information about how others behave or think can also have an impact on prejudice and its expression. For example, positive information about the racial beliefs of others produced stereotype change in message recipients \citep{stangor2001changing}, and public expression of prejudice, as well as reactions to hostile jokes were highly correlated with social approval of these expressions \citep{crandall2002social}.

When it comes to the use of normative influence to reduce uncivil behavior online, several norms were activated trough social sanctions,  peer presure and context. In a study by  \citet{munger2017tweetment} social sanctioning of racist tweets in the form of messages reminding about the human and labeling the behavior, resulted in the reduction of racist tweets, although only in the condition in which a white male with high followers  approached another white male \citep{munger2017tweetment}. 
In addition, types of statements made by people in online political discussions influenced the way others expressed their opinions \citep{price2006normative}, prior exposure to troll content was a strong predictor for trolling behavior \citep{cheng2017anyone}, and the expression of online hate-speech in \citep{alvarez2018normative} was strongly dependent on the amount of hate-speech already present in a given environment.

We have included both descriptive and prescriptive norms among normative interventions. Descriptive norms were expressed by informing the attacker about the prevalence of civil behavior in this community, and prescriptive norms were expressed by informing about the rules of  this community.

\section{Collective Intelligence and  moderation}\label{sec:collective-intelligence}

@@

Instead of using a negative motivation system, the assumption was to test a positive one—convincing verbally violent users to refrain from using cyberviolence based on civic education—peer-pressure regulation and experiential learning of a positive set of norms and empathy. Algorithms developed for the detection of personal attacks were used to monitor the activity of experimental groups and notify volunteers about all attacks generated by its members. Volunteers, after receiving a notification on Slack, could then react with an appropriate intervention. Such an approach served as a distributed bottom-up voluntary model of moderation based on collective intelligence—utilizing human + machine intelligence to develop solutions to social challenges.

Normative interventions can refer either to general social norms of civility and respect, community standards, a particular subreddit rule, or a descriptive norm among others. Empathetic interventions can refer to the emotional state of the recipient or the sender of the attack or both. They are designed to evoke an empathetic response. Here are some examples of interventions that were used during the experiment:

\begin{itemize}
  \item \emph{Insulting others is against Reddit's policy.} (normative)
  \item \emph{Hey there, we do not call each other like that here.} (normative)
  \item \emph{There is a human being on the other side who might be hurt by your words.} (empathetic)
  \item \emph{I see you are frustrated but remember the human.} (empathetic)
\end{itemize}

@@

%the reasoning behind H2
Although empathy has been found to be helpful in the reduction of aggressive behavior \citep{bjorkqvist2000social}, and the improvement of intergroup relations \citep{stephan1999role}, there are also important arguments and evidence concerning its limits. Specifically, Paul Bloom in his book \emph{Against Empathy: The Case for Rational Compassion} \citep{bloom2017against}, stresses that empathy can have a very narrow focus and can be prone to biases. For instance, it’s easier to be empathetic towards those who are close or similar to us. Since conversations on Reddit are anonymous and usually occur between strangers, we expected  empathetic interventions to  be less effective than normative interventions in curbing personal attacks.

@@

\textit{Personal attacks} are defined as any
kind of abusive remark made in relation to a person
(\textit{ad hominem}) rather than to the content of the argument
expressed by that person in a discussion. The definition of `personal
attack' subsumes the use of specific terms which compare other people to
animals or objects or make nasty insinuations without providing
evidence. Three examples of typical personal attacks are as follows:

\begin{itemize} 
\item \emph{You sound like a whiney bitch.}
\item \emph {Proof my dick in your ass u flairless fucken hoe.}
\item \emph {Go cuddle your anime pillow dipshit.} 
\end{itemize}

@@

We conducted a 6-month  experiment in a digital setting conducted on a popular Q\&A and news forum, Reddit (\url{www.reddit.com}). We formed treatment and control groups based on three main criteria:

@@

\begin{thebibliography}{115}
